---
title: "Comprehensive Analysis of Boston Airbnb Detailed Listings"
author: "Fall Foliage"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 3
    fig_width: 10
    fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  warning = FALSE, 
  message = FALSE, 
  fig.align = "center",
  fig.width = 12,
  fig.height = 8,
  fig.dpi = 300,              
  dev = "png",                
  dev.args = list(bg = "transparent"),  
  cache = TRUE
)

# Custom color palette
theme_colors <- list(
  primary = "#3498db",        
  secondary = "#1abc9c",      
  accent = "#16a085",         
  dark = "#008B8B",        
  light = "#87CEEB",         
  gradient_start = "#5DADE2", 
  gradient_end = "#138D75",   
  highlight = "#AED6F1",      
  text = "#2C3E50",          
  additional = "#3CB371"
)

# Custom ggplot theme with transparent background
theme_transparent <- function() {
  theme_minimal() +
  theme(
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    panel.grid.major = element_line(color = "grey90", linewidth = 0.5),
    panel.grid.minor = element_line(color = "grey95", linewidth = 0.25),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    plot.title = element_text(face = "bold", size = 16, color = theme_colors$text, 
                              margin = margin(b = 10)),
    plot.subtitle = element_text(size = 12, color = theme_colors$text, 
                                 margin = margin(b = 15)),
    axis.title = element_text(size = 12, color = theme_colors$text),
    axis.text = element_text(size = 10, color = theme_colors$text),
    legend.text = element_text(size = 10, color = theme_colors$text),
    legend.title = element_text(size = 11, face = "bold", color = theme_colors$text),
    strip.text = element_text(size = 11, face = "bold", color = theme_colors$text),
    strip.background = element_rect(fill = "transparent", color = NA)
  )
}
```

```{r libraries, message=FALSE, warning=FALSE}
# Core data manipulation
library(readr)
library(glue)
library(tidyr)
library(dplyr)
library(stringr)
library(lubridate)
library(purrr)
library(mgcv)


# Visualization
library(ggplot2)
library(gridExtra)
library(grid)
library(RColorBrewer)
library(viridis)
library(scales)

# Mapping and geographic analysis
library(leaflet)
library(sf)
library(ggmap)

# Statistical analysis
library(corrplot)
library(skimr)
library(cluster)
library(factoextra)
library(NbClust)

# Time series
library(xts)
library(zoo)
library(forecast)

# Text analysis (for amenities)
library(stringr)

```

# 1. Data Overview

## 1.1 Data Loading

```{r load_data}
# Load the detailed listings dataset
df_raw <- read_csv("data/listings_detailed.csv", 
                   na = c("", "NA", "N/A", "null"),
                   show_col_types = FALSE)

# Initial dataset dimensions
cat("=== DATASET OVERVIEW ===\n")
cat("Total Observations:", nrow(df_raw), "\n")
cat("Total Variables:", ncol(df_raw), "\n")
cat("\nColumn Names:\n")
cat(paste(colnames(df_raw), collapse = ", "), "\n")
```

This dataset comes from Airbnb’s public listings for Boston (as compiled by Inside Airbnb) and reflects the state of the market from June to September 2025. Each row corresponds to a single listing and includes host, property, pricing, review and availability information used later for segmentation and performance analysis.

## 1.2 Missing Data Analysis

```{r missing_data_analysis}
# Comprehensive missing data analysis
missing_analysis <- df_raw |>
  summarise(across(everything(), ~sum(is.na(.)))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "missing_count") |>
  mutate(missing_pct = round(missing_count / nrow(df_raw) * 100, 2)) |>
  filter(missing_count > 0) |>
  arrange(desc(missing_pct))

cat("=== MISSING DATA SUMMARY ===\n")
cat("Variables with missing data:", nrow(missing_analysis), "out of", ncol(df_raw), "\n\n")
print(head(missing_analysis, 20))

# Visualize missing data pattern
missing_data <-
  ggplot(missing_analysis |> head(15),
         aes(x = reorder(variable, missing_pct), y = missing_pct)) +
    geom_bar(stat = "identity", fill = theme_colors$primary, alpha = 0.8) +
    coord_flip() +
    labs(title = "Missing Data Percentage by Variable (Top 15)",
         x = "Variable",
         y = "Missing Data (%)") +
    theme_transparent() +
    theme(
      plot.title = element_text(face = "bold", color = theme_colors$text),
      panel.grid.minor = element_blank()
    )

print(missing_data)

```

Variables that were entirely missing (e.g., neighbourhood_group_cleansed, calendar_updated) were excluded from modeling and, where necessary, reconstructed from related fields.


# 2. Exploratory Data Analysis

## 2.1 Price Cleaning

```{r clean_price}
# Clean price column (remove $ and convert to numeric)
df <- df_raw |>
  mutate(
    price_clean = as.numeric(str_remove_all(price, "[$,\\s]")),
   price_category = case_when(
   price_clean < 100 ~ "Budget",
   price_clean < 170 ~ "Economy",
   price_clean < 250 ~ "Mid-range",
   price_clean < 350 ~ "Premium",
   TRUE ~ "Luxury"
)
,
    price_category = factor(price_category, 
                           levels = c("Budget", "Economy", "Mid-range", "Premium", "Luxury"))
  )

# Remove extreme outliers (top 0.5% and bottom 0.5%)
price_q01 <- quantile(df$price_clean, 0.005, na.rm = TRUE)
price_q995 <- quantile(df$price_clean, 0.995, na.rm = TRUE)

df <- df |>
  mutate(price_clean = ifelse(price_clean < price_q01 | price_clean > price_q995, 
                               NA, price_clean))

cat("Price range after cleaning: $", min(df$price_clean, na.rm = TRUE), 
    " - $", max(df$price_clean, na.rm = TRUE), "\n")
cat("Price outliers removed (0.5% tails):", 
    sum(df_raw$price_clean < price_q01 | df_raw$price_clean > price_q995, na.rm = TRUE), 
    "observations\n")
cat("Key Percentiles")
quantile(df$price_clean, probs = seq(0, 1, 0.1), na.rm = TRUE)
```

Although a 0.5% tail-trimming rule was applied to remove potential price-entry errors, no observations fell outside the statistical bounds. This indicates that the extremely high nightly prices (e.g., \$40,000–$50,000) are genuine rather than numerical anomalies. Because such ultra-luxury listings represent a meaningful market segment rather than noise, we deliberately keep them for further analysis and allow the clustering algorithm to treat them as a distinct pricing subgroup.

## 2.2 Host Response Rate and Acceptance Rate Cleaning

```{r clean_host_metrics}
# Clean host response and acceptance rates (convert percentages)
df <- df |>
  mutate(
    host_response_rate_num = as.numeric(str_remove_all(host_response_rate, "[%]")) / 100,
    host_acceptance_rate_num = as.numeric(str_remove_all(host_acceptance_rate, "[%]")) / 100,
    host_is_superhost_bool = ifelse(host_is_superhost == "TRUE", TRUE, FALSE)
  )

cat("Host response rate range:", 
    round(min(df$host_response_rate_num, na.rm = TRUE) * 100, 1), "% -",
    round(max(df$host_response_rate_num, na.rm = TRUE) * 100, 1), "%\n")
cat("Superhosts:", sum(df$host_is_superhost_bool, na.rm = TRUE), 
    "(", round(mean(df$host_is_superhost_bool, na.rm = TRUE) * 100, 1), "% )\n")
```

Almost a third of all hosts are Superhosts, meaning they are top-rated, experienced host for providing exceptional hospitality, meeting strict criteria like high response rates (90%+), low cancellation rates (<1%), high overall ratings (4.8+), and hosting a minimum number of stays/nights, earning perks like travel coupons and increased guest trust.

## 2.3 Date Column Processing

```{r process_dates}
# Convert date columns
df <- df |>
  mutate(
    host_since_date = as.Date(host_since),
    first_review_date = as.Date(first_review),
    last_review_date = as.Date(last_review),
    last_scraped_date = as.Date(last_scraped),
    # Calculate host experience in days
    host_experience_days = as.numeric(last_scraped_date - host_since_date),
    host_experience_years = host_experience_days / 365,
    # Time since first review
    days_since_first_review = as.numeric(last_scraped_date - first_review_date),
    # Time since last review
    days_since_last_review = as.numeric(last_scraped_date - last_review_date)
  )

cat("Host experience range:", round(min(df$host_experience_years, na.rm = TRUE), 1), 
    "-", round(max(df$host_experience_years, na.rm = TRUE), 1), "years\n")
```

## 2.4 Amenities Feature Engineering & Analysis

```{r process_amenities}
# Count amenities and identify key amenities
df <- df |>
  mutate(
    amenity_count = str_count(amenities, ",") + 1,
    has_wifi = str_detect(amenities, "(?i)wifi|(?i)internet"),
    has_kitchen = str_detect(amenities, "(?i)kitchen"),
    has_heating = str_detect(amenities, "(?i)heating"),
    has_ac = str_detect(amenities, "(?i)air conditioning|(?i)ac"),
    has_parking = str_detect(amenities, "(?i)parking|(?i)free parking"),
    has_washer = str_detect(amenities, "(?i)washer"),
    has_dryer = str_detect(amenities, "(?i)dryer"),
    has_tv = str_detect(amenities, "(?i)tv|(?i)television"),
    has_workspace = str_detect(amenities, "(?i)workspace|(?i)laptop"),
    has_gym = str_detect(amenities, "(?i)gym|(?i)fitness"),
    premium_amenity_score = has_wifi + has_kitchen + has_heating + has_ac + 
                           has_parking + has_washer + has_dryer + has_tv + 
                           has_workspace + has_gym
  )

amenities <- tibble(
  amenity = c("wifi", "kitchen", "heating", "ac", "parking", "washer", "dryer", "tv", "workspace", "gym"),
  listings_with_amenity = c(
    sum(df$has_wifi, na.rm = TRUE),
    sum(df$has_kitchen, na.rm = TRUE),
    sum(df$has_heating, na.rm = TRUE),
    sum(df$has_ac, na.rm = TRUE),
    sum(df$has_parking, na.rm = TRUE),
    sum(df$has_washer, na.rm = TRUE),
    sum(df$has_dryer, na.rm = TRUE),
    sum(df$has_tv, na.rm = TRUE),
    sum(df$has_workspace, na.rm = TRUE),
    sum(df$has_gym, na.rm = TRUE)
  ),
  pct_of_listings_with_amenity = c(
    round(mean(df$has_wifi, na.rm = TRUE) * 100, 2),
    round(mean(df$has_kitchen, na.rm = TRUE) * 100, 2),
    round(mean(df$has_heating, na.rm = TRUE) * 100, 2),
    round(mean(df$has_ac, na.rm = TRUE) * 100, 2),
    round(mean(df$has_parking, na.rm = TRUE) * 100, 2),
    round(mean(df$has_washer, na.rm = TRUE) * 100, 2),
    round(mean(df$has_dryer, na.rm = TRUE) * 100, 2),
    round(mean(df$has_tv, na.rm = TRUE) * 100, 2),
    round(mean(df$has_workspace, na.rm = TRUE) * 100, 2),
    round(mean(df$has_gym, na.rm = TRUE) * 100, 2)
  )
)

amenities

amenities_distribution <- amenities |>
  ggplot(aes(x = reorder(amenity, -pct_of_listings_with_amenity), y = pct_of_listings_with_amenity, fill = amenity)) +
  geom_bar(stat = "identity", fill = theme_colors$primary) +
  geom_text(aes(label = paste0(round(pct_of_listings_with_amenity, 1), "%")),
            vjust = -0.4, size = 3) +
  labs(title = "Distribution of Popular Amenities",
       x = "Amenity",
       y = "Percentage of Listings with Amenity") +
  theme_transparent() +
  theme(plot.title = element_text(face = "bold", size = 14, color = theme_colors$text),
        legend.position = "none",
        axis.text.x = element_text(angle = 15, hjust = 1),
        panel.grid.minor = element_blank())

amenities_distribution
```

The selected amenities focus on features that are both frequently present and plausibly monetizable (e.g., WiFi, kitchen, climate control, parking, in-unit laundry, workspace, gym). This compact set avoids overfitting to very rare amenities while still capturing a meaningful notion of “amenity richness” that is interpretable for hosts and guests.

## 2.5 Revenue and Performance Metrics

```{r create_performance_metrics}
# Create comprehensive performance metrics

review_subscores <- as.matrix(df[, c("review_scores_accuracy", "review_scores_cleanliness",
                                      "review_scores_checkin", "review_scores_communication",
                                      "review_scores_location", "review_scores_value")])


review_sub_mean <- rowMeans(review_subscores, na.rm = TRUE)
review_sub_mean[is.nan(review_sub_mean)] <- NA_real_


df <- df |>
  mutate(
    avg_review_score = ifelse(is.na(review_scores_rating),
                              review_sub_mean,
                              review_scores_rating),
    revenue_l365d = ifelse(is.na(estimated_revenue_l365d), 
                           price_clean * (365 - availability_365), 
                           estimated_revenue_l365d),
    revenue_per_bed = revenue_l365d / ifelse(beds > 0, beds, 1),
    revenue_per_bedroom = revenue_l365d / ifelse(bedrooms > 0, bedrooms, 1),
    occupancy_rate = ifelse(is.na(estimated_occupancy_l365d),
                            (365 - availability_365) / 365,
                            estimated_occupancy_l365d / 100),
    review_activity_score = reviews_per_month * avg_review_score,
    price_per_person = price_clean / accommodates,
    listing_efficiency = revenue_l365d / (price_clean + 1)  # Revenue per unit price
  )

df <- df |>
  mutate(
    reviews_per_month = ifelse(is.na(reviews_per_month), 0, reviews_per_month),
    number_of_reviews = ifelse(is.na(number_of_reviews), 0, number_of_reviews),
    occupancy_rate = ifelse(is.na(occupancy_rate) | occupancy_rate < 0, 0, 
                           ifelse(occupancy_rate > 1, 1, occupancy_rate))
  )

cat("Revenue metrics created successfully.\n")
cat("Average annual revenue: $", round(mean(df$revenue_l365d, na.rm = TRUE), 0), "\n")
cat("Average occupancy rate:", round(mean(df$occupancy_rate, na.rm = TRUE) * 100, 1), "%\n")
```

When Airbnb’s estimated revenue or occupancy figures were missing, we approximated them using nightly price and availability under the assumption of a constant price over the year. This introduces a mild optimistic bias for highly seasonal listings but provides a consistent, transparent way to recover performance metrics for a large share of the dataset.

## 2.6 Neighbourhood Grouping

```{r neighborhood_grouping}
# Load neighbourhood boundaries GeoJSON for mapping (used in geographic sections)
neighbourhoods_geo <- st_read("data/neighbourhoods.geojson", quiet = TRUE)
cat("\nLoaded", nrow(neighbourhoods_geo), "neighbourhood boundaries from GeoJSON\n")
```

## 2.7 Geographic Distribution Overview

```{r geographic_overview}
# Basic geographic summary
cat("=== GEOGRAPHIC COVERAGE ===\n")
cat("Latitude range:", round(min(df$latitude, na.rm = TRUE), 4), "-", 
    round(max(df$latitude, na.rm = TRUE), 4), "\n")
cat("Longitude range:", round(min(df$longitude, na.rm = TRUE), 4), "-", 
    round(max(df$longitude, na.rm = TRUE), 4), "\n")
cat("\nNeighbourhoods covered:", length(unique(df$neighbourhood_cleansed)), "\n")


df_map <- df |>
  filter(!is.na(latitude), !is.na(longitude), !is.na(price_clean),
         latitude > 42.2 & latitude < 42.4,
         longitude > -71.2 & longitude < -70.9)


df_map_sample <- df_map |>
  sample_n(min(2000, nrow(df_map)))


price_pal <- colorNumeric(
  palette = colorRampPalette(c(theme_colors$light, theme_colors$primary, 
                               theme_colors$accent, theme_colors$dark))(100),
  domain = df_map_sample$price_clean
)

price_range <- range(df_map_sample$price_clean, na.rm = TRUE)

visible_prices <- df_map_sample$price_clean[!is.na(df_map_sample$longitude)]

price_pal <- colorNumeric(
  palette = colorRampPalette(c(
    theme_colors$light,
    theme_colors$primary,
    theme_colors$accent,
    theme_colors$dark
  ))(100),
  domain = price_range
)

# Map 1: Price Distribution with Neighbourhood Boundaries
map_price <- leaflet() |>
  addProviderTiles(providers$CartoDB.Positron, options = providerTileOptions(opacity = 0.9)) |>
  addPolygons(
    data = neighbourhoods_geo,
    fillColor = "transparent",
    fillOpacity = 0,
    color = theme_colors$text,
    weight = 1.5,
    opacity = 0.6,
    label = ~neighbourhood,
    labelOptions = labelOptions(noHide = FALSE, textOnly = FALSE)
  ) |>
  addCircleMarkers(
    data = df_map_sample,
    lng = ~longitude,
    lat = ~latitude,
    radius = 4,
    color = ~price_pal(price_clean),
    fillColor = ~price_pal(price_clean),
    fillOpacity = 0.6,
    stroke = TRUE,
    weight = 0.5,
    popup = ~paste(
      "<b>Price:</b> $", round(price_clean, 0), "<br>",
      "<b>Room Type:</b>", room_type, "<br>",
      "<b>Accommodates:</b>", accommodates, "<br>",
      "<b>Neighbourhood:</b>", neighbourhood_cleansed, "<br>",
      "<b>Reviews:</b>", number_of_reviews, "<br>",
      "<b>Rating:</b>", round(review_scores_rating, 1), "/5.0"
    ),
    clusterOptions = markerClusterOptions(
      spiderfyOnMaxZoom = TRUE,
      showCoverageOnHover = TRUE,
      zoomToBoundsOnClick = TRUE
    )
  ) |>
  leaflet::addLegend(
    "bottomright",
    pal = price_pal,
    values = price_range,
    title = "Price ($)",
    opacity = 0.8
  )

map_price
```

## 2.8 Correlation Analysis

```{r correlation_analysis}
# Select numeric variables for correlation
numeric_vars <- df |>
  select(price_clean, accommodates, bedrooms, beds, bathrooms, 
         number_of_reviews, reviews_per_month, review_scores_rating,
         host_response_rate_num, host_listings_count,
         availability_365, occupancy_rate, revenue_l365d,
         amenity_count, premium_amenity_score,
         host_experience_years) |>
  drop_na() 


cor_matrix <- cor(numeric_vars)


corrplot(cor_matrix, 
         method = "color",
         type = "upper",
         order = "hclust",
         tl.col = theme_colors$text,
         tl.srt = 45,
         tl.cex = 0.7,
         addCoef.col = "black",
         number.cex = 0.6,
         col = colorRampPalette(c(theme_colors$dark, "white", theme_colors$primary))(200),
         title = "Correlation Matrix: Key Numeric Variables",
         mar = c(0, 0, 2, 0))


cat("\n=== STRONG CORRELATIONS (|r| > 0.5) ===\n")
cor_pairs <- expand.grid(row = rownames(cor_matrix), col = colnames(cor_matrix)) |>
  mutate(cor_value = as.vector(cor_matrix)) |>
  filter(row != col, abs(cor_value) > 0.5) |>
  arrange(desc(abs(cor_value))) |>
  distinct(cor_value, .keep_all = TRUE) |>
  head(10)

print(cor_pairs)
```




# 3. Host Analysis

## 3.1 Superhosts vs. Regular Hosts Performance

```{r superhost_analysis}
# Compare superhosts vs regular hosts
host_comparison <- df |>
  filter(!is.na(host_is_superhost_bool)) |>
  group_by(host_is_superhost_bool) |>
  summarise(
    n_listings = n(),
    avg_rating = mean(review_scores_rating, na.rm = TRUE),
    avg_revenue = mean(revenue_l365d, na.rm = TRUE),
    avg_occupancy = mean(occupancy_rate, na.rm = TRUE) * 100,
    avg_response_rate = mean(host_response_rate_num, na.rm = TRUE) * 100,
    pct_entire_home = mean(room_type == "Entire home/apt", na.rm = TRUE) * 100,
    .groups = "drop"
  ) |>
  mutate(host_type = ifelse(host_is_superhost_bool, "Superhost", "Regular Host"))

cat("=== SUPERHOST VS REGULAR HOST COMPARISON ===\n")
print(host_comparison)

# Visualization
p1 <- host_comparison |>
  select(host_type, avg_rating, avg_revenue, avg_occupancy) |>
  pivot_longer(cols = -host_type, names_to = "metric", values_to = "value") |>
  mutate(metric = str_remove_all(metric, "avg_") |>
                    str_replace_all("_", " ") |>
                    str_to_title(),
         metric = ifelse(metric == "Revenue", "Revenue ($)", 
                        ifelse(metric == "Occupancy", "Occupancy (%)",
                              metric))) |>
  ggplot(aes(x = metric, y = value, fill = host_type)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.8) +
  scale_fill_manual(values = c("Superhost" = theme_colors$accent,
                               "Regular Host" = theme_colors$primary)) +
  facet_wrap(~metric, scales = "free", ncol = 4) +
  labs(title = "Superhost vs Regular Host Performance",
       x = "",
       y = "Value",
       fill = "Host Type") +
  theme_transparent() +
  theme(plot.title = element_text(face = "bold", size = 14, color = theme_colors$text),
        legend.position = "bottom",
        axis.text.x = element_blank(),
        strip.text = element_text(face = "bold"))

print(p1)
```

## 3.2 Average Revenue by Host Response Rate

```{r host_type_revenue}
df_clean <- df |>
  filter(!is.na(host_response_rate_num),
         !is.na(revenue_l365d),
         host_response_rate_num >= 0,
         host_response_rate_num <= 1,
         !is.na(host_is_superhost_bool)) |>
  mutate(
    host_type = if_else(host_is_superhost_bool, "Superhost", "Regular host")
  )

df_buckets_type <- df_clean |>
  mutate(response_bucket = cut(
    host_response_rate_num,
    breaks = c(0, 0.5, 0.7, 0.9, 0.99, 1),
    labels = c("0–50%", "50–70%", "70–90%", "90–99%", "100%"),
    include.lowest = TRUE
  )) |>
  group_by(response_bucket, host_type) |>
  summarise(
    mean_revenue = mean(revenue_l365d, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) |>
  filter(!is.na(response_bucket))

p2 <- ggplot(df_buckets_type,
       aes(x = response_bucket, y = mean_revenue, fill = host_type)) +
  geom_col(position = "dodge", alpha = 0.85, color = "white") +
  scale_y_continuous(labels = scales::dollar_format()) +
  scale_fill_manual(values = c(
    "Regular host" = theme_colors$primary,
    "Superhost"    = theme_colors$accent
  )) +
  labs(
    title = "Average Annual Revenue by Response Rate and Host Type",
    x = "Host Response Rate",
    y = "Average Annual Revenue ($)",
    fill = "Host Type"
  ) +
  theme_transparent() +
  theme(
    axis.text.x = element_text(angle = 15, hjust = 1),
    legend.position = "bottom"
  )

print(p2)
```



## 3.3 Average Revenue by Host Portfolio

```{r portfolio_vs_listing_revenue_host_type}
host_portfolio_summary <- df |>
  filter(!is.na(host_listings_count),
         host_listings_count > 0,
         !is.na(revenue_l365d),
         !is.na(host_is_superhost_bool)) |>
  mutate(
    portfolio_group = case_when(
      host_listings_count == 1        ~ "Single",
      host_listings_count <= 3        ~ "Small (2–3)",
      host_listings_count <= 10       ~ "Medium (4–10)",
      host_listings_count <= 50       ~ "Large (11–50)",
      TRUE                            ~ "Very Large (50+)"
    ),
    portfolio_group = factor(
      portfolio_group,
      levels = c("Single", "Small (2–3)", "Medium (4–10)",
                 "Large (11–50)", "Very Large (50+)")
    ),
    host_type = if_else(host_is_superhost_bool, "Superhost", "Regular host")
  ) |>
  group_by(portfolio_group, host_type) |>
  summarise(
    mean_rev   = mean(revenue_l365d, na.rm = TRUE),
    n_listings = n(),
    .groups    = "drop"
  )

p3 <- ggplot(host_portfolio_summary,
       aes(x = portfolio_group, y = mean_rev, fill = host_type)) +
  geom_col(position = "dodge", alpha = 0.85, color = "white") +
  scale_y_continuous(labels = scales::dollar_format()) +
  scale_fill_manual(values = c(
    "Regular host" = theme_colors$primary,
    "Superhost"    = theme_colors$accent
  )) +
  labs(
    title    = "Average Revenue per Listing by Portfolio Size and Host Type",
    x        = "Portfolio Size",
    y        = "Avg Annual Revenue ($)",
    fill     = "Host Type",
    subtitle = "Comparing regular hosts and Superhosts at each portfolio size"
  ) +
  theme_transparent() +
  theme(
    legend.position = "bottom",
    axis.text.x     = element_text(angle = 15, hjust = 1)
  )

print(p3)
```



## 3.4 Annual Revenue Time Series by Host Experience

```{r time_series_experience}
host_exp_perf <- df |>
filter(!is.na(host_experience_years),
host_experience_years >= 0,
host_experience_years <= 15)  # trim extreme values

p4 <- ggplot(host_exp_perf, aes(x = host_experience_years,
                          y = revenue_l365d)) +
  geom_point(alpha = 0.15, color = theme_colors$primary, size = 1) +
  geom_smooth(method = "loess", color = theme_colors$accent, 
              linewidth = 1.4, se = TRUE) +
  scale_y_log10(labels = scales::dollar_format(),
                breaks = c(1000, 5000, 10000, 25000, 50000, 
                           100000, 250000, 500000, 1000000)) +
  labs(
    title = "Host Experience vs. Annual Revenue",
    subtitle = "Log scale reveals trend hidden by extreme outliers",
    x = "Years as Host",
    y = "Annual Revenue ($, log scale)"
  ) +
  theme_transparent()

print(p4)
```

This shows that most hosts—no matter how long they've been hosting—earn roughly similar annual revenue.

# 4. Property Analysis

## 4.1 Room and Property Type Analysis

```{r room_type}
# Room type distribution
room_type_dist <- df |>
  select(room_type) |>
  filter(!is.na(room_type)) |>
  count(room_type) |>
  mutate(pct_of_listings = round(n/sum(n) * 100, 2)) |>
  arrange(desc(pct_of_listings))

room_type_dist

room_type_plot <-
  ggplot(room_type_dist, aes(x = reorder(room_type, -pct_of_listings), y = pct_of_listings, fill = room_type)) +
  geom_col(width = 0.8, color = "white", linewidth = 0.5) + 
  geom_text(aes(label = paste0(round(pct_of_listings, 2), "%")),
            position = position_stack(vjust = 0.8), size = 4) +
  scale_fill_manual(values = c(theme_colors$primary, theme_colors$secondary, theme_colors$light, theme_colors$dark)) +
  labs(title = "Distribution of Room Types",
       x = "Room Type", y = "Percentage of Listings") +
  theme_transparent() +
  theme(plot.title = element_text(face = "bold", size = 14, color = theme_colors$text),
        legend.position = "none",
        axis.text.x = element_text(angle = 15, hjust = 1),
        panel.grid.minor = element_blank())
```
```{r property_type}
# Property type distribution
unique(df$property_type)
df$property_type <- case_when(
  str_detect(df$property_type, "rental unit") ~ "rental unit",
  str_detect(df$property_type, "condo") ~ "condo",
  str_detect(df$property_type, "home") ~ "home",
  str_detect(df$property_type, "townhouse") ~ "townhouse",
  str_detect(df$property_type, "guesthouse") ~ "guesthouse",
  str_detect(df$property_type, "serviced apartment") ~ "serviced apartment",
  str_detect(df$property_type, "hostel") ~ "hostel",
  str_detect(df$property_type, "hotel") ~ "hotel",
  str_detect(df$property_type, "guest suite") ~ "guest suite",
  TRUE ~ "other"
)

property_type_dist <- df |>
  select(property_type) |>
  filter(!is.na(property_type)) |>
  count(property_type) |>
  mutate(pct_of_listings = round(n/sum(n) * 100, 2)) |>
  arrange(desc(pct_of_listings))

property_type_dist

property_type_plot <-
  ggplot(property_type_dist, aes(x = reorder(property_type, pct_of_listings), y = pct_of_listings, fill = property_type)) +
  geom_col(width = 0.8, color = "white", linewidth = 0.5) + 
  geom_text(aes(label = paste0(round(pct_of_listings, 4), "%")),
            position = position_stack(vjust = 0.8), size = 4) +
  scale_fill_manual(values = c(theme_colors$primary, theme_colors$secondary, theme_colors$light, theme_colors$dark, theme_colors$accent, theme_colors$gradient_start, theme_colors$gradient_end, theme_colors$highlight, theme_colors$text, theme_colors$additional)) +
  coord_flip() +
  labs(title = "Distribution of Property Types",
       x = "Room Type", y = "Percentage of Listings") +
  theme_transparent() +
  theme(plot.title = element_text(face = "bold", size = 14, color = theme_colors$text),
        legend.position = "none",
        axis.text.x = element_text(angle = 15, hjust = 1),
        panel.grid.minor = element_blank())
```
```{r room_property_type_plots, fig.width=15, fig.height=10}
grid.arrange(room_type_plot, property_type_plot, ncol=2)
```
```{r prop_room_type_values, fig.width=20}
review_value_by_prop_and_room_type <- df |>
  select(property_type, room_type, review_scores_value) |>
  filter(!is.na(review_scores_value)) |>
  group_by(property_type, room_type) |>
  summarize(num_listings = n(),
            avg_value = round(mean(review_scores_value), 2)) |>
  filter(num_listings >= 50) |>
  arrange(desc(avg_value))

review_value_by_prop_and_room_type

ggplot(review_value_by_prop_and_room_type, aes(x = reorder(property_type, avg_value), y = room_type, fill = avg_value)) +
  geom_tile(color = "white", 
            linewidth = 0) +
  scale_fill_gradient(low = theme_colors$gradient_start, 
                      high = theme_colors$gradient_end) +
  coord_fixed() +
  labs(
    title = "Average Review Score Value by Property and Room Type",
    x = "Room Type",
    y = "Property Type",
    fill = "Avg Review Score Value"
  ) +
  theme_transparent() +
  theme(plot.title = element_text(face = "bold", size = 12, color = theme_colors$text),
        panel.grid.minor = element_blank())
```

## 4.2 Property Location Distributions

```{r locations_by_listings}

top10_neighbourhoods_by_listings <- df |>
  select(neighbourhood_cleansed) |>
  count(neighbourhood_cleansed) |>
  mutate(pct = round(n/sum(n) * 100, 2)) |>
  arrange(desc(pct)) |>
  slice_head(n = 10)

top10_neighbourhoods_by_listings

p1 <- top10_neighbourhoods_by_listings |>
  ggplot(aes(x = reorder(neighbourhood_cleansed, pct), 
             y = pct, 
             fill = pct)) +
  geom_bar(stat = "identity", 
           color = "white", 
           alpha = 0.8, 
           linewidth = 0.5) +
  geom_text(aes(label = paste0(round(pct, 2), "%")),
            position = position_stack(vjust = 0.9), size = 4) +
  coord_flip() +
  scale_fill_gradient(low = theme_colors$light, 
                      high = theme_colors$dark) +
  labs(title = "Top 10 Neighbourhoods by Number of Listings",
       x = "Neighbourhood", 
       y = "Percentage of Listings") +
  theme_transparent() +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold",
                              size = 12,
                              color = theme_colors$text),
    panel.grid.minor = element_blank()
  )
```
```{r locations_by_avg_num_of_reviews}

top10_neighbourhoods_by_avg_num_of_reviews <- df |>
  filter(!is.na(number_of_reviews)) |>
  group_by(neighbourhood_cleansed) |>
  summarize(avg_num_reviews = round(mean(number_of_reviews), 0)) |>
  arrange(desc(avg_num_reviews)) |>
  head(10)  

top10_neighbourhoods_by_avg_num_of_reviews

p2 <- top10_neighbourhoods_by_avg_num_of_reviews |>
  ggplot(aes(x = reorder(neighbourhood_cleansed, avg_num_reviews), 
                              y = avg_num_reviews, 
                              fill = avg_num_reviews)) +
  geom_bar(stat = "identity", 
           alpha = 0.8, 
           color = "white", 
           linewidth = 0.5) +
  geom_text(aes(label = avg_num_reviews),
            position = position_stack(vjust = 0.9), size = 4) +
  coord_flip() +
  scale_fill_gradient(low = theme_colors$light, 
                      high = theme_colors$dark) +
  labs(title = "Top 10 Neighbourhoods by Average Number of Reviews Per Listing",
       x = "Neighborhood",
       y = "Average Number of Reviews Per Listing",
       fill = "Avg Reviews") +
  theme_transparent() +
  theme(plot.title = element_text(face = "bold", size = 12, color = theme_colors$text),
        panel.grid.minor = element_blank(),
        legend.position = "none")

```
```{r locations_by_avg_num_of_monthly_reviews}

top10_neighbourhoods_by_avg_num_of_monthly_reviews <- df |>
  filter(!is.na(reviews_per_month)) |>
  group_by(neighbourhood_cleansed) |>
  summarize(avg_monthly_reviews = round(mean(reviews_per_month), 2)) |>
  arrange(desc(avg_monthly_reviews)) |>
  head(10)  

top10_neighbourhoods_by_avg_num_of_monthly_reviews

p3 <- top10_neighbourhoods_by_avg_num_of_monthly_reviews |>
  ggplot(aes(x = reorder(neighbourhood_cleansed, avg_monthly_reviews), 
                              y = avg_monthly_reviews, 
                              fill = avg_monthly_reviews)) +
  geom_bar(stat = "identity", 
           alpha = 0.8, color = "white", 
           linewidth = 0.5) +
  geom_text(aes(label = avg_monthly_reviews),
            position = position_stack(vjust = 0.9), size = 4) +
  coord_flip() +
  scale_fill_gradient(low = theme_colors$light, 
                      high = theme_colors$dark) +
  labs(title = "Top 10 Neighbourhoods by Average Number of Monthly Reviews Per Listing",
       x = "Neighborhood",
       y = "Average Number of Monthly Reviews Per Listing",
       fill = "Avg Monthly Reviews") +
  theme_transparent() +
  theme(plot.title = element_text(face = "bold", size = 12, color = theme_colors$text),
        panel.grid.minor = element_blank(),
        legend.position = "none")

```


```{r location_scores}
top10_neighbourhoods_by_location_scores <- df |>
  filter(!is.na(review_scores_location)) |>
  group_by(neighbourhood_cleansed) |>
  summarize(
    n = n(),
    avg_location_score = round(mean(review_scores_location), 2)) |>
  arrange(desc(avg_location_score)) |>
  filter(n >= 30) |>
  head(10)
  
top10_neighbourhoods_by_location_scores

p4 <- top10_neighbourhoods_by_location_scores |>
  ggplot(aes(x = reorder(neighbourhood_cleansed, avg_location_score), 
                              y = avg_location_score, 
                              fill = avg_location_score)) +
  geom_bar(stat = "identity", 
           alpha = 0.8, color = "white", 
           linewidth = 0.5) +
  geom_text(aes(label = avg_location_score),
            position = position_stack(vjust = 0.9), size = 4) +
  coord_flip() +
  scale_fill_gradient(low = theme_colors$light, 
                      high = theme_colors$dark) +
  labs(title = "Top 10 Neighbourhoods by Average Location Score",
       x = "Neighborhood",
       y = "Average Location Score",
       fill = "Avg Location Score") +
  theme_transparent() +
  theme(plot.title = element_text(face = "bold", size = 12, color = theme_colors$text),
        panel.grid.minor = element_blank(),
        legend.position = "none")
```
```{r plot_location_distributions, fig.width=15, fig.height=20}
grid.arrange(p1, p2, p3, p4, ncol = 1)
```

## 4.3 Property Type and Amenities

```{r property_and_amenities, fig.width=15}
# Property type and number of amenities

property_amenities <- df |>
  select(property_type, amenity_count, review_scores_value) |>
  filter(!is.na(review_scores_value)) |>
  group_by(property_type) |>
  summarize(
    num_listings = n(),
    avg_num_amenities = round(mean(amenity_count), 0),
    avg_review_score_val = round(mean(review_scores_value), 2)
  ) |>
  filter(num_listings >= 50)
    
property_amenities

property_amenities_plot <-
  ggplot(property_amenities, 
         aes(x = reorder(property_type, -avg_review_score_val), 
             y = avg_num_amenities, 
             fill = avg_review_score_val)) +
  geom_col(width = 0.8, 
           color = "white", 
           linewidth = 0.5) + 
  geom_text(aes(label = avg_num_amenities),
            position = position_stack(vjust = 0.9), 
            size = 4) +
  scale_fill_gradient(low = theme_colors$light, 
                      high = theme_colors$dark) +
  labs(title = "Property Types by Average Amenity Count and Average Review Score Value",
       x = "Property Type", 
       y = "Amenity Count",
       fill = "Average Review Score Value") +
  theme_transparent() +
  theme(plot.title = element_text(face = "bold", size = 14, color = theme_colors$text),
        axis.text.x = element_text(angle = 15, hjust = 1),
        panel.grid.minor = element_blank())

property_amenities_plot
```


 


# 5. Price and Revenue Analysis

## 5.1 Univariate Analysis - Price Distribution

```{r price_distribution}
# Price distribution analysis
p1 <- df |>
  filter(!is.na(price_clean)) |>
  ggplot(aes(x = price_clean)) +
  geom_histogram(bins = 80, fill = theme_colors$primary, alpha = 0.7, 
                 color = "white", linewidth = 0.3) +
  scale_x_continuous(labels = dollar_format(), 
                     breaks = seq(0, 1000, 100),
                     limits = c(0, 500)) +
  labs(title = "Price Distribution (Limited to $500 for Clarity)",
       x = "Price per Night ($)",
       y = "Frequency",
       subtitle = paste("Mean: $", round(mean(df$price_clean, na.rm = TRUE), 2),
                       "| Median: $", round(median(df$price_clean, na.rm = TRUE), 2))) +
  theme_transparent() +
  theme(plot.title = element_text(face = "bold", size = 14, color = theme_colors$text),
        plot.subtitle = element_text(color = theme_colors$text),
        panel.grid.minor = element_blank())

p2 <- df |>
  filter(!is.na(price_clean)) |>
  ggplot(aes(y = price_clean, x = "")) +
  geom_violin(fill = theme_colors$secondary, alpha = 0.6, color = theme_colors$dark) +
  geom_boxplot(width = 0.2, fill = theme_colors$accent, alpha = 0.8, 
               outlier.alpha = 0.3, outlier.size = 0.5) +
  scale_y_continuous(labels = dollar_format(), limits = c(0, 400)) +
  labs(title = "Price Distribution (Violin Plot)",
       y = "Price per Night ($)",
       x = "") +
  theme_transparent() +
  theme(plot.title = element_text(face = "bold", size = 14, color = theme_colors$text),
        axis.text.x = element_blank(),
        panel.grid.minor = element_blank())

p3 <- df |>
  filter(!is.na(price_category)) |>
  count(price_category) |>
  mutate(pct = n / sum(n) * 100) |>
  ggplot(aes(x = price_category, y = pct, fill = price_category)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "white", linewidth = 0.5) +
  scale_fill_manual(values = c("Budget" = theme_colors$light,
                               "Economy" = theme_colors$primary,
                               "Mid-range" = theme_colors$secondary,
                               "Premium" = theme_colors$accent,
                               "Luxury" = theme_colors$dark)) +
  geom_text(aes(label = paste0(round(pct, 1), "%")), 
            vjust = -0.5, fontface = "bold", color = theme_colors$text) +
  labs(title = "Price Category Distribution",
       x = "Price Category",
       y = "Percentage (%)") +
  theme_transparent() +
  theme(plot.title = element_text(face = "bold", size = 14, color = theme_colors$text),
        legend.position = "none",
        panel.grid.minor = element_blank())

plot_price <- grid.arrange(p1, p2, p3, ncol = 3, widths = c(1.2, 0.8, 1))
print(plot_price)

```
 
Overall, the pricing analysis suggests that Boston’s short-term rental market is characterized by a wide variety of offerings, ranging from affordable units to high-end luxury stays. This variability supports the use of clustering methods in later sections, as distinct pricing tiers likely reflect meaningful differences in property characteristics and host strategies.

## 5.2 Bivariate Analysis - Price vs Key Variables

```{r price_bivariate}
# Price by room type
p1 <- df |>
  filter(!is.na(price_clean), !is.na(room_type)) |>
  ggplot(aes(x = room_type, y = price_clean, fill = room_type)) +
  geom_violin(alpha = 0.6, color = theme_colors$dark) +
  geom_boxplot(width = 0.2, alpha = 0.8, outlier.alpha = 0.2, outlier.size = 0.3) +
  scale_fill_manual(values = c("Entire home/apt" = theme_colors$primary,
                               "Private room" = theme_colors$secondary,
                               "Shared room" = theme_colors$accent)) +
  scale_y_continuous(labels = dollar_format(), limits = c(0, 400)) +
  labs(title = "Price Distribution by Room Type",
       x = "Room Type",
       y = "Price per Night ($)") +
  theme_transparent() +
  theme(plot.title = element_text(face = "bold", size = 14, color = theme_colors$text),
        legend.position = "none",
        axis.text.x = element_text(angle = 15, hjust = 1),
        panel.grid.minor = element_blank())

# Price by accommodates
p2 <- df |>
  filter(!is.na(price_clean), !is.na(accommodates), accommodates <= 8) |>
  group_by(accommodates) |>
  summarise(
    mean_price = mean(price_clean, na.rm = TRUE),
    median_price = median(price_clean, na.rm = TRUE),
    n = n()
  ) |>
  ggplot(aes(x = accommodates, y = mean_price)) +
  geom_line(color = theme_colors$primary, linewidth = 1.5, alpha = 0.8) +
  geom_point(aes(size = n), color = theme_colors$accent, alpha = 0.7) +
  scale_y_continuous(labels = dollar_format()) +
  scale_size_continuous(range = c(2, 8), name = "Listings") +
  labs(title = "Average Price by Accommodation Capacity",
       x = "Number of Guests",
       y = "Average Price ($)") +
  theme_transparent() +
  theme(plot.title = element_text(face = "bold", size = 14, color = theme_colors$text),
        panel.grid.minor = element_blank())

# Price by bedrooms
p3 <- df |>
  filter(!is.na(price_clean), !is.na(bedrooms), bedrooms <= 5) |>
  group_by(bedrooms) |>
  summarise(
    mean_price = mean(price_clean, na.rm = TRUE),
    median_price = median(price_clean, na.rm = TRUE),
    n = n()
  ) |>
  ggplot(aes(x = bedrooms, y = mean_price)) +
  geom_col(fill = theme_colors$secondary, alpha = 0.8, color = "white", linewidth = 0.5) +
  geom_text(aes(label = dollar(round(mean_price, 0))), 
            vjust = -0.5, fontface = "bold", color = theme_colors$text) +
  scale_y_continuous(labels = dollar_format()) +
  labs(title = "Average Price by Number of Bedrooms",
       x = "Number of Bedrooms",
       y = "Average Price ($)") +
  theme_transparent() +
  theme(plot.title = element_text(face = "bold", size = 14, color = theme_colors$text),
        panel.grid.minor = element_blank())

plot_price_2 <- grid.arrange(p1, p2, p3, ncol = 3)
print(plot_price_2)
```


* **Room type** strongly segments the market into budget (shared rooms), mid-range (private rooms), and high-value (entire homes) categories.
* **Accommodation capacity** drives prices upward, particularly for large-group listings, which form a small but highly priced niche.
* **Bedroom count** is closely associated with price increases, reinforcing its role as a key structural determinant of listing value.


## 5.3 Detailed Interactive Map with Revenue Analysis

```{r interactive_map}
# Create interactive map with revenue analysis
df_map_detailed <- df_map |>
  filter(!is.na(price_clean), !is.na(revenue_l365d)) |>
  sample_n(min(2000, nrow(df_map)))

# Color palette for revenue
revenue_pal <- colorNumeric(
  palette = colorRampPalette(c(theme_colors$light, theme_colors$primary, 
                               theme_colors$accent, theme_colors$dark))(100),
  domain = df_map_detailed$revenue_l365d
)

# Create detailed map with neighbourhood boundaries
map_revenue <- leaflet() |>
  addProviderTiles(providers$CartoDB.Positron, options = providerTileOptions(opacity = 0.9)) |>
  addPolygons(
    data = neighbourhoods_geo,
    fillColor = "transparent",
    fillOpacity = 0,
    color = theme_colors$text,
    weight = 1.5,
    opacity = 0.6,
    label = ~neighbourhood,
    labelOptions = labelOptions(noHide = FALSE, textOnly = FALSE)
  ) |>
  addCircleMarkers(
    data = df_map_detailed,
    lng = ~longitude,
    lat = ~latitude,
    radius = ~sqrt(revenue_l365d / 5000),
    color = ~revenue_pal(revenue_l365d),
    fillColor = ~revenue_pal(revenue_l365d),
    fillOpacity = 0.7,
    stroke = TRUE,
    weight = 0.5,
    popup = ~paste(
      "<b>Price:</b> $", round(price_clean, 0), "<br>",
      "<b>Annual Revenue:</b> $", round(revenue_l365d, 0), "<br>",
      "<b>Room Type:</b>", room_type, "<br>",
      "<b>Occupancy Rate:</b>", round(occupancy_rate * 100, 1), "%", "<br>",
      "<b>Neighbourhood:</b>", neighbourhood_cleansed, "<br>",
      "<b>Rating:</b>", round(review_scores_rating, 1), "/5.0"
    ),
    clusterOptions = markerClusterOptions(
      spiderfyOnMaxZoom = TRUE,
      showCoverageOnHover = TRUE,
      zoomToBoundsOnClick = TRUE
    )
  )# |>
  # leaflet::addLegend(
  #   "bottomright",
  #   pal = revenue_pal,
  #   values = df_map_detailed$revenue_l365d,
  #   title = "Annual Revenue ($)",
  #   opacity = 0.8,
  #   labFormat = labelFormat(prefix = "$", big.mark = ",")
  # )

map_revenue
```



## 5.4 Revenue Optimization Analysis

```{r revenue_analysis}
# Analyze revenue drivers
revenue_analysis <- df |>
  filter(!is.na(revenue_l365d), !is.na(price_clean), revenue_l365d > 0) |>
  mutate(
    price_bucket = cut(price_clean, 
                      breaks = quantile(price_clean, probs = seq(0, 1, 0.2), na.rm = TRUE),
                      include.lowest = TRUE,
                      labels = c("Q1", "Q2", "Q3", "Q4", "Q5")),
    occupancy_bucket = cut(occupancy_rate,
                          breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1),
                          include.lowest = TRUE,
                          labels = c("0-20%", "20-40%", "40-60%", "60-80%", "80-100%"))
  )

# Revenue by price and occupancy
revenue_summary <- revenue_analysis |>
  group_by(price_bucket, occupancy_bucket) |>
  summarise(
    avg_revenue = mean(revenue_l365d, na.rm = TRUE),
    median_revenue = median(revenue_l365d, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) |>
  filter(n >= 10)

p1 <- ggplot(revenue_summary, 
             aes(x = price_bucket, y = occupancy_bucket, fill = avg_revenue)) +
  geom_tile(color = "white", linewidth = 0.5) +
  scale_fill_gradient(low = theme_colors$light, high = theme_colors$dark,
                     labels = dollar_format()) +
  labs(title = "Average Annual Revenue Heatmap",
       x = "Price Quintile",
       y = "Occupancy Rate",
       fill = "Revenue ($)") +
  theme_transparent() +
  theme(plot.title = element_text(face = "bold", size = 14, color = theme_colors$text),
        panel.grid = element_blank())

print(p1)

# Optimal pricing analysis
optimal_pricing <- revenue_analysis |>
  filter(occupancy_rate >= 0.5) |>
  group_by(price_category) |>
  summarise(
    avg_revenue = mean(revenue_l365d, na.rm = TRUE),
    avg_occupancy = mean(occupancy_rate, na.rm = TRUE) * 100,
    avg_price = mean(price_clean, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) |>
  arrange(desc(avg_revenue))

cat("\n=== REVENUE BY PRICE CATEGORY (50%+ Occupancy) ===\n")
print(optimal_pricing)
```


## 5.5 Neighbourhood Analysis

```{r neighbourhood_analysis}
# Analyze by neighbourhood
neighborhood_stats <- df |>
  filter(!is.na(neighbourhood_cleansed), !is.na(price_clean)) |>
  group_by(neighbourhood_cleansed) |>
  summarise(
    n_listings = n(),
    mean_price = mean(price_clean, na.rm = TRUE),
    median_price = median(price_clean, na.rm = TRUE),
    mean_rating = mean(review_scores_rating, na.rm = TRUE),
    mean_revenue = mean(revenue_l365d, na.rm = TRUE),
    mean_occupancy = mean(occupancy_rate, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(desc(n_listings)) |>
  head(20)

# Visualize top neighborhoods
p1 <- neighborhood_stats |>
  head(15) |>
  mutate(neighbourhood = factor(neighbourhood_cleansed, 
                               levels = rev(neighbourhood_cleansed))) |>
  ggplot(aes(x = neighbourhood, y = mean_price, fill = n_listings)) +
  geom_bar(stat = "identity", alpha = 0.8) +
  scale_fill_gradient(low = theme_colors$light, high = theme_colors$dark, 
                     name = "Listings") +
  coord_flip() +
  scale_y_continuous(labels = dollar_format()) +
  labs(title = "Average Price by Top 15 Neighbourhoods",
       x = "Neighbourhood",
       y = "Average Price ($)") +
  theme_transparent() +
  theme(plot.title = element_text(face = "bold", size = 14, color = theme_colors$text),
        panel.grid.minor = element_blank())

p2 <- neighborhood_stats |>
  head(15) |>
  mutate(neighbourhood = factor(neighbourhood_cleansed,
                               levels = rev(neighbourhood_cleansed))) |>
  ggplot(aes(x = neighbourhood, y = mean_revenue, fill = mean_occupancy)) +
  geom_bar(stat = "identity", alpha = 0.8) +
  scale_fill_gradient(low = theme_colors$light, high = theme_colors$dark,
                     labels = percent_format(), name = "Occupancy") +
  coord_flip() +
  scale_y_continuous(labels = dollar_format()) +
  labs(title = "Average Annual Revenue by Top 15 Neighbourhoods",
       x = "Neighbourhood",
       y = "Average Revenue ($)") +
  theme_transparent() +
  theme(plot.title = element_text(face = "bold", size = 14, color = theme_colors$text),
        panel.grid.minor = element_blank())

neighbourhood_plot <- grid.arrange(p1, p2, ncol = 2)

print(neighbourhood_plot)
```


## 5.6 Marginal Revenue per Additional Amenity

```{r test}
# Basic cleaning & derived columns
df2 <- df |>
  mutate(
    revenue = as.numeric(revenue_l365d),
    price = as.numeric(price_clean),
    occupancy = as.numeric(occupancy_rate),
    avg_review = as.numeric(avg_review_score),      # or review_scores_rating
    beds = as.numeric(beds),
    bedrooms = as.numeric(bedrooms),
    price_per_person = as.numeric(price_per_person),
    revenue_per_bed = as.numeric(revenue_per_bed),
    has_wifi = if_else(has_wifi %in% c(TRUE, "t", "TRUE", 1), 1, 0),
    host_experience_years = as.numeric(host_experience_years)
  ) |>
  # remove impossible / NA revenues and coordinates outside expected range
  filter(!is.na(revenue) & revenue >= 0)

glimpse(df2 |> select(revenue, price, occupancy, beds, bedrooms, avg_review, revenue_per_bed))

summary_stats <- df2 |>
  summarize(
    n = n(),
    mean_revenue = mean(revenue, na.rm=TRUE),
    median_revenue = median(revenue, na.rm=TRUE),
    sd_revenue = sd(revenue, na.rm=TRUE),
    p5 = quantile(revenue, .05, na.rm=TRUE),
    p25 = quantile(revenue, .25, na.rm=TRUE),
    p75 = quantile(revenue, .75, na.rm=TRUE),
    p95 = quantile(revenue, .95, na.rm=TRUE),
    max_revenue = max(revenue, na.rm=TRUE)
  )
print(summary_stats)
```

```{r eff}
r_am_roi <- df |>
  filter(!is.na(revenue_l365d), !is.na(amenity_count),
         revenue_l365d <= 250000, amenity_count <= 40) |>
  group_by(amenity_count) |>
  summarise(
    mean_rev = mean(revenue_l365d, na.rm = TRUE),
    n = n()
  ) |>
  mutate(
    marginal_rev = mean_rev - lag(mean_rev)
  ) |>
  ggplot(aes(x = amenity_count, y = marginal_rev)) +
  geom_col(fill = theme_colors$accent, alpha = 0.8) +
  geom_hline(yintercept = 0, color = theme_colors$dark, linewidth = 0.6) +
  labs(
    title = "Marginal Revenue per Additional Amenity",
    x = "Amenity Count",
    y = "Marginal Revenue Change ($)"
  ) +
  theme_transparent()

r_am_roi
```


## 5.7 Revenue Surface: Interaction of Price and Occupancy

```{r rev_heatmap}
df_rev_surface <- df |>
  filter(!is.na(price_clean), !is.na(occupancy_rate),
         !is.na(revenue_l365d), revenue_l365d <= 250000) |>
  mutate(
    price_bin = cut(price_clean, breaks = seq(0, 600, 50), include.lowest = TRUE),
    occ_bin = cut(occupancy_rate, breaks = seq(0, 1, 0.1), include.lowest = TRUE)
  ) |>
  group_by(price_bin, occ_bin) |>
  summarise(mean_rev = mean(revenue_l365d, na.rm = TRUE)) |> 
  ungroup()

r_heat <- df_rev_surface |>
  ggplot(aes(x = price_bin, y = occ_bin, fill = mean_rev)) +
  geom_tile() +
  scale_fill_gradient(low = theme_colors$light, high = theme_colors$dark,
                      labels = dollar_format()) +
  labs(
    title = "Revenue Surface: Interaction of Price and Occupancy",
    x = "Price per Night (Binned)",
    y = "Occupancy Rate (Binned)",
    fill = "Mean Revenue"
  ) +
  theme_transparent() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

r_heat
```


## 5.8 Revenue Elasticity

```{r}
# ------- PRICE ELASTICITY -------

df_el <- df |>
  filter(
    !is.na(price_clean),
    !is.na(revenue_l365d),
    is.finite(price_clean),
    is.finite(revenue_l365d),
    revenue_l365d <= 250000,
    revenue_l365d > 0,
    price_clean > 0,
    price_clean <= 5000
  )

# GAM on log-log scale
gam_price <- gam(log(revenue_l365d) ~ s(log(price_clean)), data = df_el)

# Grid over log(price)
lp_seq <- seq(
  from = log(min(df_el$price_clean, na.rm = TRUE)),
  to   = log(max(df_el$price_clean, na.rm = TRUE)),
  length.out = 300
)

grid_price <- data.frame(price_clean = exp(lp_seq))

# Fitted smooth on link scale (log revenue)
f_hat <- predict(gam_price, newdata = grid_price, type = "link")

# Central finite differences: d log(rev) / d log(price)
n <- length(lp_seq)
deriv_lp <- rep(NA_real_, n)
deriv_lp[2:(n - 1)] <- (f_hat[3:n] - f_hat[1:(n - 2)]) / (lp_seq[3:n] - lp_seq[1:(n - 2)])

elasticity_price <- tibble(
  price      = grid_price$price_clean,
  elasticity = deriv_lp
)

r_el_price <- elasticity_price |>
  ggplot(aes(x = price, y = elasticity)) +
  geom_line(color = theme_colors$primary, linewidth = 1.4) +
  geom_hline(yintercept = 0, color = theme_colors$dark) +
  labs(
    title = "Price Elasticity of Annual Revenue",
    subtitle = "Elasticity ≈ d log(Revenue) / d log(Price)",
    x = "Price ($)",
    y = "Elasticity"
  ) +
  theme_transparent()


# ------- OCCUPANCY ELASTICITY -------

df_el_occ <- df |>
  filter(
    !is.na(occupancy_rate),
    !is.na(revenue_l365d),
    revenue_l365d <= 250000,
    revenue_l365d > 0,
    occupancy_rate > 0,
    occupancy_rate < 1
  )

gam_occ <- gam(log(revenue_l365d) ~ s(log(occupancy_rate)), data = df_el_occ)

lo_seq <- seq(
  from = log(min(df_el_occ$occupancy_rate, na.rm = TRUE)),
  to   = log(max(df_el_occ$occupancy_rate, na.rm = TRUE)),
  length.out = 300
)

grid_occ <- data.frame(occupancy_rate = exp(lo_seq))

g_hat <- predict(gam_occ, newdata = grid_occ, type = "link")

n2 <- length(lo_seq)
deriv_lo <- rep(NA_real_, n2)
deriv_lo[2:(n2 - 1)] <- (g_hat[3:n2] - g_hat[1:(n2 - 2)]) / (lo_seq[3:n2] - lo_seq[1:(n2 - 2)])

elasticity_occ <- tibble(
  occ        = grid_occ$occupancy_rate,
  elasticity = deriv_lo
)

r_el_occ <- elasticity_occ |>
  ggplot(aes(x = occ, y = elasticity)) +
  geom_line(color = theme_colors$accent, linewidth = 1.4) +
  geom_hline(yintercept = 0, color = theme_colors$dark) +
  labs(
    title = "Occupancy Elasticity of Annual Revenue",
    subtitle = "Elasticity ≈ d log(Revenue) / d log(Occupancy)",
    x = "Occupancy rate",
    y = "Elasticity"
  ) +
  theme_transparent()

# Show both plots
grid.arrange(r_el_price, r_el_occ, ncol = 2)



```



## 5.9 Revenue Misc

```{r rev test setup}
# ---------- Global cleaning & derived columns ----------
df_clean <- df |>
  mutate(
    revenue = as.numeric(revenue_l365d),
    price = as.numeric(price_clean),
    occupancy = as.numeric(occupancy_rate),
    avg_review = coalesce(as.numeric(avg_review_score), as.numeric(review_scores_rating)),
    beds = as.numeric(beds),
    bedrooms = as.numeric(bedrooms),
    amenity_count = as.numeric(amenity_count),
    host_years = as.numeric(host_experience_years),
    is_superhost = as.integer(host_is_superhost_bool),
    listings_owned = as.numeric(calculated_host_listings_count),
    listing_eff = as.numeric(listing_efficiency)
  ) |>
  # remove obviously bad coords / values and keep reasonable ranges
  filter(is.finite(revenue), is.finite(price), is.finite(occupancy)) |>
  filter(revenue >= 0, price > 0, occupancy >= 0, occupancy <= 1) |>
  # cap extreme revenues for stability (keeps original for reporting if needed)
  mutate(revenue_capped = pmin(revenue, 250000))

# A quick summary
df_clean |> summarise(n = n(), mean_rev = mean(revenue), median_rev = median(revenue), p95 = quantile(revenue, .95))

# ---------- Helpers ----------
# backtransform function (for log models)
inv_log10 <- function(x) 10^x - 1
```

```{r rev test}
minn_df <- df_clean |>
  filter(minimum_nights <= 30) |>
  group_by(minimum_nights) |>
  summarise(mean_occ = mean(occupancy, na.rm=TRUE), mean_rev = mean(revenue, na.rm=TRUE), n = n())

p_min <- ggplot(minn_df, aes(x = minimum_nights, y = mean_rev)) +
  geom_line(color = theme_colors$primary, size=1.2) +
  geom_point() + scale_y_continuous(labels = dollar_format()) +
  labs(title = "Mean Revenue vs Minimum Nights", x = "Minimum Nights", y = "Mean Revenue ($)") +
  theme_transparent()
p_min

df_eff <- df_clean |>
  filter(availability_365 >= 0 & availability_365 <= 365 & price <= 5000) |>
  mutate(avail_nights = 365 - availability_365,
         rev_per_avail = ifelse(avail_nights > 0, revenue / avail_nights, NA))

# Plot efficiency vs price
p_eff <- ggplot(df_eff |> filter(!is.na(rev_per_avail) & rev_per_avail < 2000), aes(x = price, y = rev_per_avail)) +
  geom_point(alpha = 0.25) + geom_smooth(method="loess", se=FALSE, color = theme_colors$accent) +
  labs(title = "Revenue per Available Night vs Price", x = "Price ($)", y = "Revenue / Available Night ($)") +
  theme_transparent()
p_eff

# Model with host features
df_host <- df_clean |> filter(revenue > 0 & revenue <= 250000)
host_mod <- lm(log10(revenue) ~ host_years + is_superhost + host_response_rate_num + listings_owned + avg_review, data = df_host)
summary(host_mod)

# Simple visual
p_host <- df_host |> filter(host_years <= 20) |>
  ggplot(aes(x = host_years, y = revenue)) +
  geom_point(alpha=0.2) + geom_smooth(method="loess", se=FALSE, color = theme_colors$secondary) +
  scale_y_continuous(labels = dollar_format()) +
  labs(title = "Revenue vs Host Experience (years)") +
  theme_transparent()
p_host

nb_summary <- df_clean |>
  group_by(neighbourhood_cleansed) |>
  summarise(nb_median = median(revenue, na.rm=TRUE), n = n())

bench <- df_clean |> left_join(nb_summary, by = "neighbourhood_cleansed") |>
  mutate(rel_to_nb = revenue / nb_median) |>
  arrange(neighbourhood_cleansed, desc(rel_to_nb)) 

# Underperformers: rel_to_nb < 0.6 in neighborhoods with n >= 10
underperf <- bench |> filter(n >= 10, rel_to_nb < 0.6)
underperf |> select(id, neighbourhood_cleansed, revenue, nb_median, rel_to_nb) |> head(20)

sup_demand <- df_clean |>
  group_by(neighbourhood_cleansed) |>
  summarise(median_price = median(price, na.rm=TRUE),
            median_occ = median(occupancy, na.rm=TRUE),
            n = n()) |>
  mutate(supply_demand_index = median_price * median_occ) |>
  arrange(desc(supply_demand_index))

head(sup_demand, 15)
```

* If revenue falls as minimum_nights increases, recommend lowering minimum nights to increase bookings.
* If revenue increases with minimum nights (and occupancy doesn’t drop too much), longer minimums may be beneficial for host stability.

* Low rev_per_avail suggests either low price, low occupancy, or both — investigate listings with good location but low efficiency.
* Recommend: improve photos, reviews, adjust price, add amenity investments.

# 6. Clustering Analysis

```{r clustering_prep}
# Prepare data for clustering
clustering_data <- df |>
  filter(
    !is.na(price_clean),
    !is.na(latitude),
    !is.na(longitude),
    !is.na(accommodates),
    !is.na(bedrooms),
    !is.na(review_scores_rating),
    !is.na(revenue_l365d)
  ) |>
  mutate(
    # Normalize key features for clustering
    price_scaled = scale(price_clean)[,1],
    revenue_scaled = scale(revenue_l365d)[,1],
    rating_scaled = scale(review_scores_rating)[,1],
    accommodates_scaled = scale(accommodates)[,1],
    latitude_scaled = scale(latitude)[,1],
    longitude_scaled = scale(longitude)[,1],
    occupancy_scaled = scale(occupancy_rate)[,1],
    amenity_scaled = scale(premium_amenity_score)[,1]
  ) |>
  select(id, price_clean, revenue_l365d, review_scores_rating, accommodates,
         bedrooms, latitude, longitude, occupancy_rate, premium_amenity_score,
         room_type, neighbourhood_cleansed,
         price_scaled, revenue_scaled, rating_scaled, accommodates_scaled,
         latitude_scaled, longitude_scaled, occupancy_scaled, amenity_scaled)

cat("Prepared", nrow(clustering_data), "observations for clustering\n")
cat("Features selected:", 
    paste(c("Price", "Revenue", "Rating", "Capacity", "Location", 
            "Occupancy", "Amenities"), collapse = ", "), "\n")
```

```{r optimal_clusters}
# Select features for clustering
cluster_features <- clustering_data |>
  select(price_scaled, revenue_scaled, rating_scaled, 
         accommodates_scaled, occupancy_scaled, amenity_scaled)

# Method 1: Elbow method
set.seed(123)
wss <- map_dbl(1:10, function(k) {
  kmeans(cluster_features, centers = k, nstart = 25, iter.max = 50)$tot.withinss
})

elbow_df <- data.frame(k = 1:10, wss = wss)

p1 <- ggplot(elbow_df, aes(x = k, y = wss)) +
  geom_line(color = theme_colors$primary, linewidth = 1.2) +
  geom_point(color = theme_colors$accent, size = 3) +
  scale_x_continuous(breaks = 1:10) +
  labs(title = "Elbow Method for Optimal Clusters",
       x = "Number of Clusters (k)",
       y = "Within Sum of Squares",
       subtitle = "Look for the 'elbow' in the curve") +
  theme_transparent() +
  theme(plot.title = element_text(face = "bold", size = 14, color = theme_colors$text),
        panel.grid.minor = element_blank())

# Method 2: Silhouette method
silhouette_scores <- map_dbl(2:10, function(k) {
  km <- kmeans(cluster_features, centers = k, nstart = 25, iter.max = 50)
  ss <- silhouette(km$cluster, dist(cluster_features))
  mean(ss[, 3])
})

sil_df <- data.frame(k = 2:10, silhouette = silhouette_scores)

p2 <- ggplot(sil_df, aes(x = k, y = silhouette)) +
  geom_line(color = theme_colors$secondary, linewidth = 1.2) +
  geom_point(color = theme_colors$accent, size = 3) +
  scale_x_continuous(breaks = 2:10) +
  labs(title = "Silhouette Method for Optimal Clusters",
       x = "Number of Clusters (k)",
       y = "Average Silhouette Width",
       subtitle = "Higher values indicate better clustering") +
  theme_transparent() +
  theme(plot.title = element_text(face = "bold", size = 14, color = theme_colors$text),
        panel.grid.minor = element_blank())

cluster_plot <-grid.arrange(p1, p2, ncol = 2)
print(cluster_plot)

# Determine optimal k
optimal_k <- sil_df$k[which.max(sil_df$silhouette)]
cat("\n=== OPTIMAL NUMBER OF CLUSTERS ===\n")
cat("Based on silhouette method: k =", optimal_k, "\n")
cat("Average silhouette width:", round(max(sil_df$silhouette), 3), "\n")
```

## K-Means Clustering

```{r kmeans_clustering}
# Perform K-means clustering with optimal k
set.seed(123)
kmeans_result <- kmeans(cluster_features, centers = optimal_k, nstart = 50, iter.max = 100)

# Add cluster assignments
clustering_data$cluster <- as.factor(kmeans_result$cluster)

cat("=== CLUSTER ASSIGNMENTS ===\n")
print(table(clustering_data$cluster))

# Prepare data for geographic cluster visualization
df_clustered_geo <- df |>
  left_join(
    clustering_data |> select(id, cluster),
    by = "id"
  ) |>
  filter(!is.na(cluster), !is.na(latitude), !is.na(longitude),
         latitude > 42.2 & latitude < 42.4,
         longitude > -71.2 & longitude < -70.9)

# Sample for performance
df_clustered_geo_sample <- df_clustered_geo |>
  sample_n(min(2000, nrow(df_clustered_geo)))

# Create distinct cluster color palette - using high contrast, easily distinguishable colors
# Define colors for each cluster (1-5)
cluster_colors_list <- c("#3498db", "#2ecc71", "#f39c12", "#e74c3c", "#9b59b6")
names(cluster_colors_list) <- as.character(1:5)

# Ensure cluster is character for consistent color mapping
df_clustered_geo_sample$cluster_char <- as.character(df_clustered_geo_sample$cluster)

# Create cluster color palette with all possible cluster values
all_clusters <- sort(unique(df_clustered_geo_sample$cluster_char))
cluster_pal <- colorFactor(
  palette = cluster_colors_list[all_clusters],
  domain = all_clusters
)

# Visualize clusters on geographic map - WITHOUT clustering to show individual colors
map_clusters_geo <- leaflet() |>
  addProviderTiles(providers$CartoDB.Positron, options = providerTileOptions(opacity = 0.9)) |>
  addPolygons(
    data = neighbourhoods_geo,
    fillColor = "transparent",
    fillOpacity = 0,
    color = theme_colors$text,
    weight = 1.5,
    opacity = 0.6,
    label = ~neighbourhood,
    labelOptions = labelOptions(noHide = FALSE, textOnly = FALSE)
  ) |>
  addCircleMarkers(
    data = df_clustered_geo_sample,
    lng = ~longitude,
    lat = ~latitude,
    radius = 5,
    color = "white",
    fillColor = ~cluster_pal(cluster_char),
    fillOpacity = 0.8,
    stroke = TRUE,
    weight = 1.5,
    popup = ~paste(
      "<b>Cluster:</b>", cluster, "<br>",
      "<b>Price:</b> $", round(price_clean, 0), "<br>",
      "<b>Revenue:</b> $", round(revenue_l365d, 0), "<br>",
      "<b>Rating:</b>", round(review_scores_rating, 1), "/5.0", "<br>",
      "<b>Neighbourhood:</b>", neighbourhood_cleansed, "<br>",
      "<b>Room Type:</b>", room_type, "<br>",
      "<b>Occupancy:</b>", round(occupancy_rate * 100, 1), "%"
    )
  ) |>
  leaflet::addLegend(
    "bottomright",
    pal = cluster_pal,
    values = all_clusters,
    title = paste("Cluster (k =", optimal_k, ")"),
    opacity = 0.9
  )

print(map_clusters_geo)
```

## Cluster Characterization

```{r cluster_characterization}

# Analyze cluster characteristics using ALL relevant features
cluster_profiles <- clustering_data |>
  group_by(cluster) |>
  summarise(
    n   = n(),
    pct = round(n / nrow(clustering_data) * 100, 1),

    # --- RAW FEATURES ---
    avg_price        = mean(price_clean, na.rm = TRUE),
    median_price     = median(price_clean, na.rm = TRUE),
    avg_revenue      = mean(revenue_l365d, na.rm = TRUE),
    avg_rating       = mean(review_scores_rating, na.rm = TRUE),
    avg_occupancy    = mean(occupancy_rate, na.rm = TRUE) * 100,   # as %
    avg_accommodates = mean(accommodates, na.rm = TRUE),
    avg_bedrooms     = mean(bedrooms, na.rm = TRUE),
    avg_amenities    = mean(premium_amenity_score, na.rm = TRUE),
    pct_entire_home  = mean(room_type == "Entire home/apt", na.rm = TRUE) * 100,

    # --- RAW LOCATION ---
    avg_latitude     = mean(latitude, na.rm = TRUE),
    avg_longitude    = mean(longitude, na.rm = TRUE),

    # --- SCALED FEATURES (CENTROIDS IN FEATURE SPACE) ---
    price_scaled_mean        = mean(price_scaled, na.rm = TRUE),
    revenue_scaled_mean      = mean(revenue_scaled, na.rm = TRUE),
    rating_scaled_mean       = mean(rating_scaled, na.rm = TRUE),
    accommodates_scaled_mean = mean(accommodates_scaled, na.rm = TRUE),
    latitude_scaled_mean     = mean(latitude_scaled, na.rm = TRUE),
    longitude_scaled_mean    = mean(longitude_scaled, na.rm = TRUE),
    occupancy_scaled_mean    = mean(occupancy_scaled, na.rm = TRUE),
    amenity_scaled_mean      = mean(amenity_scaled, na.rm = TRUE),

    # --- DOMINANT CATEGORICAL FEATURES ---
    main_room_type      = names(which.max(table(room_type))),

    .groups = "drop"
  ) |>
  arrange(desc(avg_revenue))

cat("=== CLUSTER PROFILES ===\n")
print(
  cluster_profiles |>
    mutate(
      avg_price     = round(avg_price, 2),
      median_price  = round(median_price, 2),
      avg_revenue   = round(avg_revenue, 0),
      avg_rating    = round(avg_rating, 2),
      avg_occupancy = round(avg_occupancy, 1),
      avg_accommodates = round(avg_accommodates, 1),
      avg_bedrooms  = round(avg_bedrooms, 1),
      avg_amenities = round(avg_amenities, 1),
      pct_entire_home = round(pct_entire_home, 1)
    )
)

# Use the same cluster colors as the maps for consistency
cluster_colors_plot <- c(
  "1" = "#3498db",      # Bright Blue
  "2" = "#2ecc71",      # Green
  "3" = "#f39c12",      # Orange
  "4" = "#e74c3c",      # Red
  "5" = "#9b59b6"       # Purple
)

# Visualize key numeric characteristics (for the report)
p1 <- cluster_profiles |>
  select(cluster, avg_price, avg_revenue, avg_rating, avg_occupancy) |>
  pivot_longer(cols = c(avg_price, avg_revenue, avg_rating, avg_occupancy),
               names_to = "metric", values_to = "value") |>
  mutate(
    metric = metric |>
      str_remove_all("avg_") |>
      str_replace_all("_", " ") |>
      str_to_title(),
    cluster = as.character(cluster)
  ) |>
  ggplot(aes(x = cluster, y = value, fill = cluster)) +
  geom_bar(stat = "identity", alpha = 0.8, position = "dodge") +
  facet_wrap(~ metric, scales = "free_y", ncol = 2) +
  scale_fill_manual(values = cluster_colors_plot) +
  labs(title = "Cluster Characteristics Comparison",
       x = "Cluster",
       y = "Value",
       fill = "Cluster") +
  theme_transparent() +
  theme(
    plot.title   = element_text(face = "bold", size = 14, color = theme_colors$text),
    legend.position = "none",
    strip.text   = element_text(face = "bold")
  )

print(p1)

# -------------------------
# Name the clusters based on ALL characteristics
# -------------------------

cluster_names <- cluster_profiles |>
  mutate(
    cluster = as.integer(cluster),

    # PRICE: uses price_scaled_mean (price_scaled) and avg_price
    price_segment = case_when(
      price_scaled_mean >=  1.0 ~ "Luxury",
      price_scaled_mean >=  0.5 ~ "Upscale",
      price_scaled_mean <= -1.0 ~ "Budget",
      price_scaled_mean <= -0.5 ~ "Economy",
      TRUE                      ~ "Mid-Priced"
    ),

    # REVENUE: uses revenue_scaled_mean
    revenue_segment = case_when(
      revenue_scaled_mean >=  0.8 ~ "High-Revenue",
      revenue_scaled_mean <= -0.8 ~ "Low-Revenue",
      TRUE                        ~ "Mid-Revenue"
    ),

    # RATING: uses rating_scaled_mean (review_scores_rating / rating_scaled)
    rating_segment = case_when(
      rating_scaled_mean >=  0.7 ~ "Top-Rated",
      rating_scaled_mean <= -0.7 ~ "Lower-Rated",
      TRUE                       ~ "Mid-Rated"
    ),

    # OCCUPANCY: uses occupancy_scaled_mean (occupancy_rate / occupancy_scaled)
    occupancy_segment = case_when(
      occupancy_scaled_mean >=  0.8 ~ "High-Occupancy",
      occupancy_scaled_mean <= -0.8 ~ "Low-Occupancy",
      TRUE                          ~ "Mid-Occupancy"
    ),

    # AMENITIES: uses amenity_scaled_mean (premium_amenity_score / amenity_scaled)
    amenity_segment = case_when(
      amenity_scaled_mean >=  0.6 ~ "Amenity-Rich",
      amenity_scaled_mean <= -0.6 ~ "Amenity-Light",
      TRUE                        ~ "Standard Amenities"
    ),

    # SIZE: uses accommodates, bedrooms, and accommodates_scaled
    size_segment = case_when(
      avg_accommodates >= 6 | avg_bedrooms >= 3 ~ "Large-Group",
      avg_accommodates >= 3 | avg_bedrooms >= 2 ~ "Family-Sized",
      TRUE                                      ~ "Compact"
    ),


    # FINAL HUMAN-READABLE LABEL
    cluster_label = glue(
      "{price_segment} {revenue_segment}, {rating_segment}, ",
      "{occupancy_segment}, {amenity_segment} {size_segment} ",
      "{main_room_type}"
    )
  ) |>
  # If two clusters somehow get exactly the same label, disambiguate by cluster id
  group_by(cluster_label) |>
  mutate(
    cluster_label = if_else(
      n() > 1,
      glue("{cluster_label} (Cluster {cluster})"),
      cluster_label
    )
  ) |>
  ungroup() |>
  # Round some things for nicer printing
  mutate(
    avg_price        = round(avg_price, 2),
    median_price     = round(median_price, 2),
    avg_revenue      = round(avg_revenue, 0),
    avg_rating       = round(avg_rating, 2),
    avg_occupancy    = round(avg_occupancy, 1),
    avg_accommodates = round(avg_accommodates, 1),
    avg_bedrooms     = round(avg_bedrooms, 1),
    avg_amenities    = round(avg_amenities, 1),
    pct_entire_home  = round(pct_entire_home, 1)
  ) |>
  select(
    cluster,
    cluster_label,
    n, pct,
    avg_price, median_price, avg_revenue, avg_rating,
    avg_occupancy, avg_accommodates, avg_bedrooms, avg_amenities,
    pct_entire_home,
    main_room_type
  )

cat("\n=== CLUSTER LABELS ===\n")
print(cluster_names |> select(cluster, cluster_label, avg_price, avg_revenue))

```
| Original                                             | Cleaner Name                 |
| ---------------------------------------------------- | ---------------------------- |
| **Luxury High-Revenue Large-Group Hotel**            | **Luxury Hotel Cluster**     |
| **Mid-Priced Amenity-Rich Large-Group Entire Homes** | **Amenity-Rich Large Homes** |
| **Mid-Priced Compact Entire Homes**                  | **Standard Mid-Tier Homes**  |
| **Lower-Rated Low-Occupancy Compact Homes**          | **Underperforming Units**    |
| **Mid-Rated Low-Occupancy Compact Homes**            | **Low-Demand Homes**         |


A small number of listings (n=14) have extremely high nightly prices (\$40–50k) and revenues exceeding \$300k–$2M annually. These represent luxury hotel penthouses and group suites. Because of their magnitude, k-means correctly isolates them into a separate cluster. We retain them to preserve meaningful segmentation, but acknowledge their outlier nature.

```{r cluster names}
cluster_names <- cluster_names |>
  mutate(
    cluster_name_short = case_when(
      cluster == 3 ~ "Luxury Hotel Cluster",
      cluster == 2 ~ "Amenity-Rich Large Homes",
      cluster == 1 ~ "Standard Mid-Tier Homes",
      cluster == 5 ~ "Underperforming Units",
      cluster == 4 ~ "Low-Demand Homes",
      TRUE         ~ as.character(cluster_label)
    )
  )
cluster_names
```

## Geographic Distribution of Clusters

```{r cluster_geography, fig.width=14, fig.height=7}
# Attach cluster ids to full df
df_clustered <- df |>
  left_join(
    clustering_data |> select(id, cluster),
    by = "id"
  ) |>
  filter(!is.na(cluster))

# --- Top neighbourhoods by cluster ---
cluster_neighborhood <- df_clustered |>
  filter(!is.na(cluster), !is.na(neighbourhood_cleansed)) |>
  count(cluster, neighbourhood_cleansed) |>
  # make cluster numeric so it joins cleanly with cluster_names
  mutate(cluster = as.integer(as.character(cluster))) |>
  left_join(
    cluster_names |> select(cluster, cluster_name_short),
    by = "cluster"
  ) |>
  group_by(neighbourhood_cleansed) |>
  mutate(pct = n / sum(n) * 100) |>
  ungroup() |>
  group_by(cluster, cluster_name_short) |>
  slice_max(pct, n = 5, with_ties = FALSE) |>
  ungroup()

cat("\n=== TOP NEIGHBOURHOODS BY CLUSTER (share of listings in each neighbourhood) ===\n")
print(cluster_neighborhood)

# --- Plot: top neighbourhoods for each cluster ---
cluster_neighborhood_plot <- cluster_neighborhood |>
  mutate(
    neighbourhood_cleansed =
      forcats::fct_reorder(neighbourhood_cleansed, pct)
  ) |>
  ggplot(aes(x = neighbourhood_cleansed, y = pct, fill = cluster_name_short)) +
  geom_col(alpha = 0.9, show.legend = FALSE) +
  coord_flip() +
  facet_wrap(
    ~ cluster_name_short,
    scales = "free_y",
    ncol = 3,
    labeller = label_wrap_gen(width = 20)   # <-- wrap long cluster titles
  ) +
  labs(
    title = "Top Neighbourhoods for Each Cluster",
    x = "Neighbourhood",
    y = "Share of Listings in Neighbourhood (%)"
  ) +
  theme_transparent() +
  theme(
    plot.title = element_text(
      face = "bold", size = 16, color = theme_colors$text
    ),

    # --- fix truncated facet titles ---
    strip.text = element_text(
      face = "bold",
      size = 12,
      margin = margin(t = 10, r = 5, b = 10, l = 5)
    ),

    # more padding around the whole plot (prevents clipping)
    plot.margin = margin(20, 20, 20, 20)
  )

print(cluster_neighborhood_plot)
```

---

# References and Data Sources

- **Dataset**: Boston Airbnb Detailed Listings (`listings_detailed.csv`)
- **Analysis Date**: `r Sys.Date()`
- **Tools**: R (tidyverse, ggplot2, leaflet, cluster, forecast)

---

*Analysis conducted by Fall Foliage Team*